---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
actData <- read.csv(unz("activity.zip", "activity.csv"), 
                    colClasses = c("numeric", "character","integer"), header = T)
```

Parse date in the order year, month, and days 
```{r}
library(lubridate)
actData$date <- ymd(actData$date)
```

## What is mean total number of steps taken per day?
```{r}
totSteps <- tapply(actData$steps, actData$date, FUN = sum, na.rm = TRUE)
mean(totSteps)
```

Also the median could be calculated as follows

```{r}
median(totSteps)
```

For the total number of steps taken by day 
```{r}
library(plyr)
library(dplyr)
steps <- print(summarize(group_by(filter(actData, !is.na(steps)), date), steps = sum(steps)))

```
## What is the average daily activity pattern?
The histogram for the average daily activity

```{r}
library(ggplot2)
ggplot(steps, aes(x=date, y=steps)) + 
        geom_histogram(stat="identity", colour = "black", fill = "green") +
#        geom_vline(aes(xintercept=mean(, na.rm=T)),   # Ignore NA values for mean
#               color="red", linetype="dashed", size=1) +
        xlab("Dates") + ylab("Steps") + 
        labs(title= "Total numbers of Steps per day")
       
```



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
