---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
actData <- read.csv(unz("activity.zip", "activity.csv"), 
                    colClasses = c("numeric", "character","integer"), header = T)
```

Parse date in the order year, month, and days 
```{r}
library(lubridate)
actData$date <- ymd(actData$date)
```

## What is mean total number of steps taken per day?
```{r}
totSteps <- tapply(actData$steps, actData$date, FUN = sum, na.rm = TRUE)
mean(totSteps)
```

Also the median could be calculated as follows

```{r}
median(totSteps)
```

For the total number of steps taken by day 
```{r}
library(plyr)
library(dplyr)
dailySteps <- print(summarize(group_by(filter(actData, !is.na(steps)), date), dailySteps = sum(steps)))
```

The histogram for the total daily steps

```{r}
library(ggplot2)
ggplot(dailySteps, aes(x=date, y=dailySteps)) + 
        geom_histogram(stat="identity", colour = "black", fill = "green") +
        xlab("Days") + ylab("Number of Steps") + 
        labs(title= "Total number of steps per day") 
```



## What is the average daily activity pattern?
Calculating the average number of steps in each interval

```{r}
intervalAvgSteps <- print(summarize(group_by(filter(actData, !is.na(steps)), interval), intervalAvgSteps = mean(steps)))
```

The time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
plot(intervalAvgSteps, type = "l", xlab = "The 5-minute Intervals", ylab = "Avg. number of steps", main = "The average number of steps across all days")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
actData$interval[max(intervalAvgSteps[2])]
```


## Imputing missing values
Calculating the total number of missing values for all the coloumns
```{r}
totMissingSteps       <- sum(is.na(actData$steps))
totMissingDates       <- sum(is.na(actData$date))
totMissingIntervals   <- sum(is.na(actData$interval))
print(totMissingSteps + totMissingDates + totMissingIntervals)
```
Note There are only missing values in the **steps** coloumn 


####I'll use the mean of the total number of steps in an interval to fill in the NA value's places 

```{r}
newData <- mutate(group_by(actData, interval), steps = ifelse(is.na(steps), 
                                           mean(steps, na.rm=TRUE), steps))
head(newData)
```

I'll get the new steps in a separate variable 
```{r}
newSteps <- print(summarize(group_by(newData, date), newSteps = sum(steps)))
```

Histogram for the new values

```{r}
ggplot(newSteps, aes(x=date, y=newSteps)) + 
        geom_histogram(stat="identity", colour = "black", fill = "red") +
        xlab("Days") + ylab("New Number of Steps") + 
        labs(title= "The New total number of steps per day") 
```

The mean of the new days
```{r}
newDailySteps <- tapply(newData$steps, newData$date, FUN = sum)
averageNewDailySteps <- mean(newDailySteps)
head(averageNewDailySteps)
```

The median of the new days
```{r}
medianNewDailySteps <- median(newDailySteps)
head(medianNewDailySteps)
```

#### We can notice from the results that both the mean and median differ now from the initial estimates, and we can prove it

```{r}
mean   (totSteps)   == mean   (newDailySteps)
median (totSteps)   == median (newDailySteps)
```

#### We can show the impact of the imputed data using the summary subtraction of both cases [before/after] imputing the data

```{r}
summary(newDailySteps) - summary(totSteps)
```

## Are there differences in activity patterns between weekdays and weekends?

#### Creating a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
newData$whatDay <- as.factor(ifelse(weekdays(as.Date(newData$date)) %in% c("Saturday", "Sunday"), "weekend", "normalDay"))
summary(newData$whatDay)
```


####Plotting the average number of steps/interval for both of weekdays and weekends 
```{r}
for (type in c("weekend", "normalDay")) {
    whatStep <- aggregate(steps ~ interval, data = newData, 
                            subset = newData$whatDay == type, FUN = mean)
    plot(whatStep, type = "l", main = type)
}
```














